Decentralised Authoring, Annotations and Notifications for a Read-Write Web with dokieli

### 2017-04-05 icwe-2017-review-response.md

Dear Sarven,

We are very pleased to inform you that your submission entitled

Decentralised Authoring, Annotations and Notifications for a Read-Write-Web with dokieli

has been ACCEPTED as an application paper at ICWE 2017.

** CAMERA READY INSTRUCTION **

Please prepare the camera-ready version of your paper (maximum length is 12 pages) by **April 7th** by addressing the reviewer comments below and follow STRICTLY the Springer LNCS guidelines for authors at http://www.springer.de/comp/lncs/authors.html. This is a HARD deadline, no extensions will be granted. Springer may reject your paper if the manuscript is not properly formatted. The camera-ready version of your paper (sources in a zip + separate pdf) and the copyright form (in pdf, signed and scanned) should be uploaded on EasyChair.

A second email will follow with detailed instructions.

In case any further questions arise during the preparation of your camera-ready version, please do not hesitate to contact the program chairs AND the proceedings Chair (Florian Daniel, florian.daniel at polimi.it).

** CONFERENCE AUTHOR REGISTRATION **

Please do not forget that at least one author per accepted submission needs to register as a CONFERENCE AUTHOR attendee in order for the paper to appear in the proceedings. Information about the registration procedure is available on the conference Web site. If you have multiple papers accepted, each paper must have a separate CONFERENCE AUTHOR registration.

Thank you for choosing ICWE 2017; we look forward to seeing you in Rome!

Kind regards, 
ICWE 2017 Application Track Chair

----------------------- REVIEW 1 ---------------------
PAPER: 152
TITLE: Decentralised Authoring, Annotations and Notifications for a Read-Write-Web with dokieli
AUTHORS: Sarven Capadisli, Amy Guy, Ruben Verborgh, Christoph Lange, Sören Auer and Tim Berners-Lee

----------- Overall evaluation -----------
This paper presents dokieli, a decentralized web-based authoring service. 
The paper is very well-written and a pleasure to read. The service presented here is nice and useful, it defines a new way of publishing and interacting with content on the Web. There are many on-line resources that can be triggered to understand more about this service; indeed, the reader can access the github repository to gather more information and also the dokieli Website is a useful resource. Nevertheless, the paper itself does not provide many technical details and remains quite at a high level.

The contribution is well-suited for the conference and it should generate interesting discussions at the conference.

----------------------- REVIEW 2 ---------------------
PAPER: 152
TITLE: Decentralised Authoring, Annotations and Notifications for a Read-Write-Web with dokieli
AUTHORS: Sarven Capadisli, Amy Guy, Ruben Verborgh, Christoph Lange, Sören Auer and Tim Berners-Lee

----------- Overall evaluation -----------
The authors present dokieli, a  decentralised, browser-based authoring and annotation platform to create “living” documents that are interoperable and independent of dokieli. Dokieli follows standards and best practices, mainly HTML+RDFa for, Linked Data Platform for personal data storage, and Linked Data Notifications for updates. The presented platform has real potential in the context of academic/journalistic writings. The paper presents good arguments around the use of such a decentralised system to offer authors ownership and sovereignty over their data. The implementation is also sound. However, the paper lacks rigour from the perspective of comparing work against existing approaches. The authors do compare and contrast their system with other offerings; however, the research work and techniques used by other is not critically analysed (for example, there are almost no references in the main text. However, there is an excellent list of references at the end, was th!
 is intentional?).

----------------------- REVIEW 3 ---------------------
PAPER: 152
TITLE: Decentralised Authoring, Annotations and Notifications for a Read-Write-Web with dokieli
AUTHORS: Sarven Capadisli, Amy Guy, Ruben Verborgh, Christoph Lange, Sören Auer and Tim Berners-Lee

----------- Overall evaluation -----------
The authors illustrate a decentralized framework for web-based authoring and annotation service.

The paper is really well-written, organized and motivated.

The main weakness is the nature of the proposal more fitting a demonstration context rather than an application/industrial scenario.

However the contribution will leverage a fruitful discussion during the conference.

### 2017-04-05 index.tex
Update references

                            \begin{description}
  \item[Data storage independent of service providers] Users should have a choice in where they store their data and full control over it e.g. with regard to who is allowed to access it. The \empty Industrial Data Space initiative calls this ``data sovereignty''.\item[Interoperability] By allowing the application logic to be decoupled from the data, users can switch between applications and personal data storage servers, thereby avoiding a {\em vendor lock-in}. To achieve maximum interoperability, applications should conform to well-defined Web standards and protocols (rather than properietary software implementations). Dangers of data silos and some example standards to use to decentralise are given in \empty Socially-aware Cloud Storage.\item[Separation of concerns] A {\em progressive enhancement} strategy to connect the structural, presentational, and behavioural layers allows content and base functionality to be accessible through different media and devices (as described in \empty Progressive Enhancement and the Future of Web Design).\item[Accessibility] To lower the barrier for entry for all forms of participation, enhanced functionality should be accessible to users based on the capabilities of their user-agents, storage availability, network access or personal preferences (we consider this to be self-evident, and there are a plethora of Web best practices in this area).\item[Freedom of expression] Because there are no central authorities, we must assume applications follow the open-world principle, where ``any author can say anything about anything''. Identifying everything using [de]referenceable IRIs allows any distributed authoring or annotation application to reference and link to previously published content (this overlaps with Principles 1, 3 and 4 in the \empty W3C Semantic Web Activity charter.\item[Web of Trust] The Web as a collaborative medium makes it possible for people to take responsibility (or be accountable) for their contributions. It should be possible for people to publish, share, and annotate information while ensuring their provenance, authenticity and integrity [\empty Weaving a Web of Trust, \empty Trust Management for the Semantic Web, \empty Trust Networks on the Semantic Web].
    \end{description}

@@

\par \empty User Interfaces for Semantic Authoring of Textual Content gives an overview on relevant related work. A range of quality attributes such as collaboration, interoperability, and scalability, while relevant to our work, we also consider systems and tools on dimensions based on the principles that we have outlined.

@@

\par On the other hand, if desired and available, users can participate using their own profiles (WebIDs) located anywhere on the Web, and get to store and make their annotations in their own personal storage, as well as assign access controls to documents. Similarly, a decentralised communications protocol, \empty Linked Data Notifications (\empty W3C Proposed Recommendation), is used get past the limits of centralisation by enabling communication to happen across independent servers. Figure 2 depicts relations between the kinds of entities which underlay dokieli’s architecture, where nodes are under different domains and authority.

@@

\par Documents use HTML5 \empty Polyglot Markup to ensure that when served as (X)HTML respectively, they can be processed as either HTML or XML, which is useful in XML ecosystems and toolchains. 
                                    Semantics is embedded directly into human-visible prose using RDFa. The machine-readable data is thus kept in context, reusing the article’s text as literal object values, and avoiding data duplication or data ‘islands’, which can occur when other RDF serialisations are included within HTML {\tt <script>} elements.

@@

\par Several stylesheets provide alternative views for consumption (e.g., stylesheets for different media: screens, print, slideshow). dokieli’s JavaScript includes: a library for editing (\empty MediumEditor) in its {\em authoring environment}; features to {\em fetch and display statistical data} from SPARQL endpoints (\empty Sparqlines); retrieval of profile information, as well as means to {\em sign in} with \empty WebID Authentication over TLS; functionality for write-operations, which includes checking {\em authorisation} against access-control level settings on the server with the authenticated user’s WebID and personal certificate; creation and consumption of \empty Web Annotations and \empty Linked Data Notifications; and fetching information from remote articles when adding a citation.

@@

\par Annotations: A core feature to facilitate collaboration is the possibility to annotate arbitrary parts of a Web document.
                                    Users can select an entity or a span of text of interest, a context menu is presented to input their annotations along with the choice to select a license for their contribution.
                                    If the user is signed-in with their WebID, and provided they have a personal storage space, dokieli discovers this through their online profile and saves the annotations to that location.
                                    Once the annotation is submitted by the user, dokieli proceeds with three operations: 1) the annotation is requested to be saved at the user’s personal storage, and if it is access controlled, the user will be prompted to authenticate themselves against that server, before the annotation is saved and is assigned its own URL, 2) if the article or any identifiable statement or segment has its own inbox, dokieli sends a notification to the inbox indicating that an annotation was made and where its retrievable location is, with accompanying metadata like who created it, date, its license etc., 3) the annotation is fetched from its canonical location, and integrated into the article e.g., in marginalia. If the article has a reference to a public annotation service (a writeable space adhering to \empty Web Annotation Protocol), the user has the option to send a copy of the annotation there as well. In cases where the user does not want to have the canonical copy of the annotation on their server, or if a user does not have write access to a storage, they can still use this option to engage with the article.

@@

\par The W3C Working Group Note \empty Embedding Web Annotations in HTML includes examples from dokieli’s use of the Web Annotation data \empty Model and \empty Vocabulary with motivations for example for ``Lightweight, decentralised Annotation Tools''.
                                    The \empty Linked Data Notifications specifications use dokieli’s HTML+RDFa template, and the \empty Editor’s Draft showcase dokieli as a consumer of LDN and Web Annotations. The LDN \empty tests suite also uses dokieli’s templates and stylesheets.

@@

                            \begin{enumerate}
  \item Otto, B., Auer, S., et al: Industrial Data Space, TR, 2016, \empty https://www.researchgate.net/publication/299536747\_Industrial\_Data\_Space\_Digital\_Souvereignity\_Over\_Data\item Berners-Lee, T.: Socially-aware Cloud Storage, 2009, \empty https://www.w3.org/DesignIssues/CloudStorage.html\item Champeon, S.: Progressive Enhancement and the Future of Web Design, 2003, \empty http://hesketh.com/publications/progressive\_enhancement\_and\_the\_future\_of\_web\_design.html\item Koivunen, M-R., Eric Miller, E.: W3C Semantic Web Activity, 2001, \empty https://www.w3.org/2001/12/semweb-fin/w3csw\item Golbeck, J.: Weaving a Web of Trust, Science Magazine, Vol. 321, 2008, \empty https://pdfs.semanticscholar.org/5e37/daf945094ea9c9df127e06b05282e03e39bd.pdf\item Richardson, M., Agrawal, R., Domingos, P.: Trust Management for the Semantic Web, ISWC 2003, \empty https://link.springer.com/content/pdf/10.1007\%2F978-3-540-39718-2\_23.pdf\item Jennifer Golbeck, J.: Parsia, B., Hendler, J.: Trust Networks on the Semantic Web, Volume 2782 of the book series LNCS, 2003, \empty http://sir-lab.usc.edu/cs586/20151readings/w13-1.pdf\item Khalili, A., Auer, S.: User interfaces for semantic authoring of textual content: A systematic literature review, Web Semantics, Volume 22, pages 1–18 (2013), \empty http://svn.aksw.org/papers/2011/JWS\_SemanticContentAuthoring/public.pdf\item Capadisli, S., Guy, A., Lange, C., Auer, S., Deiu, A., Berners-Lee, T.: Linked Data Notifications: a resource-centric communication protocol, ESWC Linked Data, 2017, \empty http://csarven.ca/linked-data-notifications\item Capadisli, S., Guy, A.: Linked Data Notifications, W3C Proposed Recommendation, 2017, \empty https://www.w3.org/TR/ldn/\item Graff, E., Silli, L.: Polyglot Markup: A robust profile of the HTML5 vocabulary, W3C Working Group Note, 2015, \empty https://www.w3.org/TR/html-polyglot/\item Capadisli, S.: Sparqlines: SPARQL to Sparkline, ISWC SemStats, 2016, \empty http://csarven.ca/sparqlines-sparql-to-sparkline\item Story, H., Corlosquet, S., Sambra, A.: WebID Authentication over TLS, W3C Editor’s Draft, 2014\empty https://www.w3.org/2005/Incubator/webid/spec/tls/\item Sanderson, R., Ciccarese, P., Young, B.: Web Annotation Data Model, W3C Recommendation, 2017, \empty https://www.w3.org/TR/annotation-model/\item Sanderson, R.: Web Annotation Protocol, W3C Recommendation, 2017, \empty https://www.w3.org/TR/annotation-protocol/\item Cole, T., Capadisli, S., Young, B., Herman, I.: Embedding Web Annotations in HTML, W3C Working Group Note, 2017, \empty https://www.w3.org/TR/annotation-html/\item Sanderson, R., Ciccarese, P., Young, B.: Web Annotation Vocabulary, W3C Recommendation, 2017, \empty https://www.w3.org/TR/annotation-vocab/
    \end{enumerate}

>>>>

\begin{description}
  \item[Data storage independent of service providers] Users should have a choice in where they store their data and full control over it e.g. with regard to who is allowed to access it. The Industrial Data Space~\cite{ref-1} initiative calls this ``data sovereignty''.

  \item[Interoperability] By allowing the application logic to be decoupled from the data, users can switch between applications and personal data storage servers, thereby avoiding a {\em vendor lock-in}. To achieve maximum interoperability, applications should conform to well-defined Web standards and protocols (rather than properietary software implementations). Dangers of data silos and some example standards to use to decentralise are given in Socially-aware Cloud Storage~\cite{ref-2}.

  \item[Separation of concerns] A {\em progressive enhancement} strategy to connect the structural, presentational, and behavioural layers allows content and base functionality to be accessible through different media and devices (as described in Progressive Enhancement and the Future of Web Design~\cite{ref-3}).

  \item[Accessibility] To lower the barrier for entry for all forms of participation, enhanced functionality should be accessible to users based on the capabilities of their user-agents, storage availability, network access or personal preferences (we consider this to be self-evident, and there are a plethora of Web best practices in this area).

  \item[Freedom of expression] Because there are no central authorities, we must assume applications follow the open-world principle, where ``any author can say anything about anything''. Identifying everything using [de]referenceable IRIs allows any distributed authoring or annotation application to reference and link to previously published content (this overlaps with Principles 1, 3 and 4 in the W3C Semantic Web Activity~\cite{ref-4} charter.

  \item[Web of Trust] The Web as a collaborative medium makes it possible for people to take responsibility (or be accountable) for their contributions. It should be possible for people to publish, share, and annotate information while ensuring their provenance, authenticity and integrity ~\cite{ref-5, ref-6, ref-7}.
\end{description}

@@

\par
User Interfaces for Semantic Authoring of Textual Content~\cite{ref-8} gives an overview on relevant related work. A range of quality attributes such as collaboration, interoperability, and scalability, while relevant to our work, we also consider systems and tools on dimensions based on the principles that we have outlined.

@@

\par On the other hand, if desired and available, users can participate using their own profiles (WebIDs) located anywhere on the Web, and get to store and make their annotations in their own personal storage, as well as assign access controls to documents. Similarly, a decentralised communications protocol, \empty Linked Data Notifications~\cite{ref-9} (W3C Proposed Recommendation~\cite{ref-10}), is used get past the limits of centralisation by enabling communication to happen across independent servers. Figure 2 depicts relations between the kinds of entities which underlay dokieli’s architecture, where nodes are under different domains and authority.

@@

\par Documents use HTML5 Polyglot Markup~\cite{ref-11} to ensure that when served as (X)HTML respectively, they can be processed as either HTML or XML, which is useful in XML ecosystems and toolchains. Semantics is embedded directly into human-visible prose using RDFa. The machine-readable data is thus kept in context, reusing the article’s text as literal object values, and avoiding data duplication or data ‘islands’, which can occur when other RDF serialisations are included within HTML {\tt <script>} elements.

@@

\par Several stylesheets provide alternative views for consumption (e.g., stylesheets for different media: screens, print, slideshow). dokieli’s JavaScript includes: a library for editing (\empty MediumEditor) in its {\em authoring environment}; features to {\em fetch and display statistical data} from SPARQL endpoints (Sparqlines~\cite{ref-12}); retrieval of profile information, as well as means to {\em sign in} with \empty WebID Authentication over TLS~\cite{ref-13}; functionality for write-operations, which includes checking {\em authorisation} against access-control level settings on the server with the authenticated user’s WebID and personal certificate; creation and consumption of \empty Web Annotations and \empty Linked Data Notifications; and fetching information from remote articles when adding a citation.

@@

\par Annotations: A core feature to facilitate collaboration is the possibility to annotate arbitrary parts of a Web document.
                                    Users can select an entity or a span of text of interest, a context menu is presented to input their annotations along with the choice to select a license for their contribution.
                                    If the user is signed-in with their WebID, and provided they have a personal storage space, dokieli discovers this through their online profile and saves the annotations to that location.
                                    Once the annotation is submitted by the user, dokieli proceeds with three operations: 1) the annotation is requested to be saved at the user’s personal storage, and if it is access controlled, the user will be prompted to authenticate themselves against that server, before the annotation is saved and is assigned its own URL, 2) if the article or any identifiable statement or segment has its own inbox, dokieli sends a notification to the inbox indicating that an annotation was made and where its retrievable location is, with accompanying metadata like who created it, date, its license etc., 3) the annotation is fetched from its canonical location, and integrated into the article e.g., in marginalia. If the article has a reference to a public annotation service (a writeable space adhering to Web Annotation Protocol~\cite{ref-15}), the user has the option to send a copy of the annotation there as well. In cases where the user does not want to have the canonical copy of the annotation on their server, or if a user does not have write access to a storage, they can still use this option to engage with the article.

@@

\par The W3C Working Group Note Embedding Web Annotations~\cite{ref-16} in HTML includes examples from dokieli’s use of the Web Annotation data Model~\cite{ref-14} and Vocabulary~\cite{ref-17} with motivations for example for ``Lightweight, decentralised Annotation Tools''.
                                    The \empty Linked Data Notifications specifications use dokieli’s HTML+RDFa template, and the \empty Editor’s Draft showcase dokieli as a consumer of LDN and Web Annotations. The LDN \empty tests suite also uses dokieli’s templates and stylesheets.

@@

\begin{thebibliography}{99}
  \bibitem{ref-1} Otto, B., Auer, S., et al: Industrial Data Space, TR, 2016, \url{https://www.researchgate.net/publication/299536747\_Industrial\_Data\_Space\_Digital\_Souvereignity\_Over\_Data}
  \bibitem{ref-2} Berners-Lee, T.: Socially-aware Cloud Storage, 2009, \url{https://www.w3.org/DesignIssues/CloudStorage.html}
  \bibitem{ref-3} Champeon, S.: Progressive Enhancement and the Future of Web Design, 2003, \url{http://hesketh.com/publications/progressive\_enhancement\_and\_the\_future\_of\_web\_design.html}
  \bibitem{ref-4} Koivunen, M-R., Eric Miller, E.: W3C Semantic Web Activity, 2001, \url{https://www.w3.org/2001/12/semweb-fin/w3csw}
  \bibitem{ref-5} Golbeck, J.: Weaving a Web of Trust, Science Magazine, Vol. 321, 2008, \url{https://pdfs.semanticscholar.org/5e37/daf945094ea9c9df127e06b05282e03e39bd.pdf}
  \bibitem{ref-6} Richardson, M., Agrawal, R., Domingos, P.: Trust Management for the Semantic Web, ISWC 2003, \url{https://link.springer.com/content/pdf/10.1007\%2F978-3-540-39718-2\_23.pdf}
  \bibitem{ref-7} Jennifer Golbeck, J.: Parsia, B., Hendler, J.: Trust Networks on the Semantic Web, Volume 2782 of the book series LNCS, 2003, \url{http://sir-lab.usc.edu/cs586/20151readings/w13-1.pdf}
  \bibitem{ref-8} Khalili, A., Auer, S.: User interfaces for semantic authoring of textual content: A systematic literature review, Web Semantics, Volume 22, pages 1–18 (2013), \url{http://svn.aksw.org/papers/2011/JWS\_SemanticContentAuthoring/public.pdf}
  \bibitem{ref-9} Capadisli, S., Guy, A., Lange, C., Auer, S., Deiu, A., Berners-Lee, T.: Linked Data Notifications: a resource-centric communication protocol, ESWC Linked Data, 2017, \url{http://csarven.ca/linked-data-notifications}
  \bibitem{ref-10} Capadisli, S., Guy, A.: Linked Data Notifications, W3C Proposed Recommendation, 2017, \url{https://www.w3.org/TR/ldn/}
  \bibitem{ref-11} Graff, E., Silli, L.: Polyglot Markup: A robust profile of the HTML5 vocabulary, W3C Working Group Note, 2015, \url{https://www.w3.org/TR/html-polyglot/}
  \bibitem{ref-12} Capadisli, S.: Sparqlines: SPARQL to Sparkline, ISWC SemStats, 2016, \url{http://csarven.ca/sparqlines-sparql-to-sparkline}
  \bibitem{ref-13} Story, H., Corlosquet, S., Sambra, A.: WebID Authentication over TLS, W3C Editor’s Draft, 2014\url{https://www.w3.org/2005/Incubator/webid/spec/tls/}
  \bibitem{ref-14} Sanderson, R., Ciccarese, P., Young, B.: Web Annotation Data Model, W3C Recommendation, 2017, \url{https://www.w3.org/TR/annotation-model/}
  \bibitem{ref-15} Sanderson, R.: Web Annotation Protocol, W3C Recommendation, 2017, \url{https://www.w3.org/TR/annotation-protocol/}
  \bibitem{ref-16} Cole, T., Capadisli, S., Young, B., Herman, I.: Embedding Web Annotations in HTML, W3C Working Group Note, 2017, \url{https://www.w3.org/TR/annotation-html/}
  \bibitem{ref-17} Sanderson, R., Ciccarese, P., Young, B.: Web Annotation Vocabulary, W3C Recommendation, 2017, \url{https://www.w3.org/TR/annotation-vocab/}
\end{thebibliography}
